Bayesian Rate Estimation

### 2012-08-16 many.tex
erf(x/2) => erf(x/sqrt(2)) typo fixed.
Changed labels in data cumulative distribution.
Updated N-templates fitting.

For a data stream of pure noise, $d(f) = n(f)$, the SNR of a each
template follows a $N(0,1)$ distribution.  The background ranking
statistic (i.e.\ the maximum SNR over the template bank) then has a
cumulative distribution without thresholding of
\begin{equation}
  \hat{B}(x) = \left( \frac{1 + \erf\left( \frac{x}{\ilya{\sqrt{2}}}
    \right)}{2} \right)^N
\end{equation}
Imposing the threshold, $x > \xmin$, the cumulative distribution of
the background becomes
\begin{equation}
  \label{eq:analytic-background-rate}
  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{\ilya{\sqrt{2}}}
    \right) \right)^N - \left( 1 + \erf\left(
    \frac{\xmin}{\ilya{\sqrt{2}}} \right) \right)^N}{2^N - \left( 1 +
    \erf\left( \frac{\xmin}{\ilya{\sqrt{2}}} \right) \right)^N }
\end{equation}
\ilya{for $x>\xmin$, $0$ otherwise.}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{data}
  \caption{\label{fig:analytic-data-cumulative} The cumulative
    distribution of the ranking statistics for the synthetic data used
    to test the formalism on the model from \S
    \ref{sec:analytic-GW-example}.  The solid line gives the
    cumulative distribution of the synthetic data; the dashed line
    gives the theoretical cumulative distribution for the models in
    Eqs.~\eqref{eq:analytic-background-rate} and
    \eqref{eq:analytic-foreground-rate} combined with $R_f = 10.4$ and
    $R_b = 95.1$. \ilya{[I think both axis labels should say $x$
        rather than $\rho$, since that's the ranking statistic ($\max
        \rho$).]}}
\end{figure}

@@

In \S \ref{sec:analytic-GW-example} we assumed that the overlap
between \ilya{different} templates in the template bank \ilya{was
  negligible}, so the SNRs \ilya{recovered by} different templates are
independent random variables.  In fact, template banks are not
constructed in this way \citep[e.g.,][]{Owen:1998dk,Ajith:2008},
%\citep{Caudill2011,Cannon2011}, 
because signals could fall in the gaps between the non-overlapping
templates.  We can model this effect by assuming that a template bank
of $N$ actual templates will behave as if it had $N_\mathrm{eff}$
\emph{independent} templates.  \ilya{Rather than pre-computing
  $N_\mathrm{eff}$,} we can fit for \ilya{it as a} shape parameter.
That is, we assume that \ilya{$\theta = \{N_\mathrm{eff}\}$} is a
shape parameter for the background cumulative distribution:
\begin{equation}
  \hat{B}\left(x, N_\mathrm{eff}\right) = \frac{\left( 1 + \erf\left(
    \frac{x}{2} \right) \right)^{N_\mathrm{eff}} - \left( 1 +
    \erf\left( \frac{\xmin}{2} \right)
    \right)^{N_\mathrm{eff}}}{2^{N_\mathrm{eff}} - \left( 1 +
    \erf\left( \frac{\xmin}{2} \right) \right)^{N_\mathrm{eff}} }.
\end{equation}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{rates-nt}
  \caption{\label{fig:rates-nt} The foreground (solid lines) and
    background (dashed lines) rate posterior, marginalized over all
    flags and the $N_\mathrm{eff}$ parameter, for the gravitational
    wave template detection scenario with overlapping templates
    discussed in \S \ref{sec:gw-overlapping-template}.  The true
    values of the rates, $R_f = 10.4$ and $R_b=95.1$, are indicated
    with vertical lines.  The distributions are not significantly
    wider than those of Figure \ref{fig:analytic-rate-recovery}, in
    spite of the extra parameter.  \ilya{[$R_f$ seems to be more of an
        outlier than before; why was there a significant shift, given
        that $N_\mathrm{eff}$ seems fairly spot on?]}}
\end{figure}

>>>>

For a data stream of pure noise, $d(f) = n(f)$, the SNR of a each
template follows a $N(0,1)$ distribution.  The background ranking
statistic (i.e.\ the maximum SNR over the template bank) then has a
cumulative distribution without thresholding of
\begin{equation}
  \hat{B}(x) = \left( \frac{1 + \erf\left( \frac{x}{\sqrt{2}}
    \right)}{2} \right)^N
\end{equation}
Imposing the threshold, $x > \xmin$, the cumulative distribution of
the background becomes
\begin{equation}
  \label{eq:analytic-background-rate}
  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{\sqrt{2}}
    \right) \right)^N - \left( 1 + \erf\left(
    \frac{\xmin}{\sqrt{2}} \right) \right)^N}{2^N - \left( 1 +
    \erf\left( \frac{\xmin}{\sqrt{2}} \right) \right)^N }
\end{equation}
\ilya{for $x>\xmin$, $0$ otherwise.}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{data}
  \caption{\label{fig:analytic-data-cumulative} The cumulative
    distribution of the ranking statistics for the synthetic data used
    to test the formalism on the model from \S
    \ref{sec:analytic-GW-example}.  The solid line gives the
    cumulative distribution of the synthetic data; the dashed line
    gives the theoretical cumulative distribution for the models in
    Eqs.~\eqref{eq:analytic-background-rate} and
    \eqref{eq:analytic-foreground-rate} combined with $R_f = 10.4$ and
    $R_b = 95.1$.}
\end{figure}

@@

In \S \ref{sec:analytic-GW-example} we assumed that the overlap
between \ilya{different} templates in the template bank \ilya{was
  negligible}, so the SNRs \ilya{recovered by} different templates are
independent random variables.  In fact, template banks are not
constructed in this way \citep[e.g.,][]{Owen:1998dk,Ajith:2008},
%\citep{Caudill2011,Cannon2011}, 
because signals could fall in the gaps between the non-overlapping
templates.  We can model this effect by assuming that a template bank
of $N$ actual templates will behave as if it had $N_\mathrm{eff}$
\emph{independent} templates.  \ilya{Rather than pre-computing
  $N_\mathrm{eff}$,} we can fit for \ilya{it as a} shape parameter.
That is, we assume that \ilya{$\theta = \{N_\mathrm{eff}\}$} is a
shape parameter for the background cumulative distribution:
\begin{equation}
  \hat{B}\left(x, N_\mathrm{eff}\right) = \frac{\left( 1 + \erf\left(
    \frac{x}{\sqrt{2}} \right) \right)^{N_\mathrm{eff}} - \left( 1 +
    \erf\left( \frac{\xmin}{\sqrt{2}} \right)
    \right)^{N_\mathrm{eff}}}{2^{N_\mathrm{eff}} - \left( 1 +
    \erf\left( \frac{\xmin}{\sqrt{2}} \right) \right)^{N_\mathrm{eff}} }.
\end{equation}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{rates-nt}
  \caption{\label{fig:rates-nt} The foreground (solid lines) and
    background (dashed lines) rate posterior, marginalized over all
    flags and the $N_\mathrm{eff}$ parameter, for the gravitational
    wave template detection scenario with overlapping templates
    discussed in \S \ref{sec:gw-overlapping-template}.  The true
    values of the rates, $R_f = 10.4$ and $R_b=95.1$, are indicated
    with vertical lines.  The distributions are not significantly
    wider than those of Figure \ref{fig:analytic-rate-recovery}, in
    spite of the extra parameter.}
\end{figure}
